---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyEmoji

<!-- badges: start -->
[![R-CMD-check](https://github.com/PursuitOfDataScience/tidyEmoji/workflows/R-CMD-check/badge.svg)](https://github.com/PursuitOfDataScience/tidyEmoji/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/tidyEmoji)](https://CRAN.R-project.org/package=tidyEmoji)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html#maturing)
<!-- badges: end -->

tidyEmoji helps you **discover, count, categorise, sentiment-score, score
emotions and translate the emoji in any text column** — social-media posts,
product reviews, chat logs, survey responses, support tickets — and hands the
result back as tidy data frames that drop straight into a tidyverse workflow.

Unicode is awkward to work with and not every code point is an emoji, which
makes emoji statistics fiddly. tidyEmoji takes care of that, including
grapheme-aware detection so skin-tone modifiers (👍🏽) and multi-person
sequences (👨‍👩‍👧‍👦) are treated as a single emoji rather than being split apart.

## Installation

``` r
install.packages("tidyEmoji")

# development version
# install.packages("devtools")
devtools::install_github("PursuitOfDataScience/tidyEmoji")
```

## Usage

```{r}
library(tidyEmoji)
library(dplyr)

reviews <- data.frame(text = c("Best purchase ever \U0001f600\U0001f60d",
                               "It broke after a day \U0001f621",
                               "Does the job.",
                               "Wearing my mask \U0001f637\U0001f637",
                               "Shipped fast \U0001f3c1\U0001f600"))
```

### How much emoji is in the data?

```{r}
reviews %>% emoji_summary(text)        # entries with emoji vs. total
reviews %>% emoji_filter(text)         # keep only the rows that have emoji
```

### Which emoji are most common?

```{r}
reviews %>% emoji_frequency(text)      # every emoji, with name + category
reviews %>% top_n_emojis(text, n = 3)  # just the most frequent
```

### Pull the emoji out

`emoji_tokens()` gives one tidy row per emoji occurrence, with its name,
category and sentiment — ready to count, join or plot.

```{r}
reviews %>% emoji_tokens(text)
```

### Categorise and score sentiment

```{r}
reviews %>% emoji_categorize(text)     # which Unicode categories each row spans
reviews %>% emoji_sentiment(text)      # mean emoji sentiment per row (-1 to +1)
```

`emoji_sentiment()` uses the bundled **Emoji Sentiment Ranking** lexicon
(Kralj Novak et al., 2015). See the package vignette for a fuller tour.

### Score emoji emotions

`emoji_emotion()` scores each row's emoji across the eight Plutchik emotions
using the bundled EmoTag1200 lexicon (Shoeb & de Melo, 2020).

```{r}
reviews %>% emoji_emotion(text) %>% select(text, .emoji_joy, .emoji_trust,
                                           .emoji_anger, .emoji_n)
reviews %>% emoji_emotion_label(text)   # the dominant emotion per row
```

### Bring your own lexicon

The scoring machinery is pluggable: `emoji_lexicons()` lists what's available,
`register_emoji_lexicon()` adds your own, and `emoji_score()` scores a text
column against any of them — always joining through a codepoint-normalised key,
so qualified and unqualified emoji forms both match.

```{r}
my_lexicon <- data.frame(emoji = c("\U0001f600", "\U0001f621"),
                         score = c(1, -1))
reviews %>% emoji_score(text, lexicon = my_lexicon)
```

### Translate emoji to and from text

`emoji_to_text()` replaces emoji with words (handy for accessibility and NLP
preprocessing); `text_to_emoji()` is the inverse. `as_emoji_name()`,
`as_emoji_shortcode()` and `as_emoji()` are the vector-level equivalents.

```{r}
reviews %>% emoji_to_text(text, format = "name")
reviews %>% emoji_to_text(text, format = "shortcode")
```

You can also search the emoji catalogue by keyword:

```{r}
emoji_search("happy")
```

## Learn more

The [introductory
vignette](https://pursuitofdatascience.github.io/tidyEmoji/articles/introduction.html)
(`vignette("introduction", package = "tidyEmoji")`) walks through a full
analysis of a real corpus — counting, categorising, sentiment- and
emotion-scoring emoji, and plotting the results.
