---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyEmoji

<!-- badges: start -->
[![R-CMD-check](https://github.com/PursuitOfDataScience/tidyEmoji/workflows/R-CMD-check/badge.svg)](https://github.com/PursuitOfDataScience/tidyEmoji/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/tidyEmoji)](https://CRAN.R-project.org/package=tidyEmoji)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

tidyEmoji helps you **discover, count, categorise and sentiment-score the emoji
in any text column** — social-media posts, product reviews, chat logs, survey
responses, support tickets — and hands the result back as tidy data frames that
drop straight into a tidyverse workflow.

Unicode is awkward to work with and not every code point is an emoji, which
makes emoji statistics fiddly. tidyEmoji takes care of that, including
grapheme-aware detection so skin-tone modifiers (👍🏽) and multi-person
sequences (👨‍👩‍👧‍👦) are treated as a single emoji rather than being split apart.

## Installation

``` r
install.packages("tidyEmoji")

# development version
# install.packages("devtools")
devtools::install_github("PursuitOfDataScience/tidyEmoji")
```

## Usage

```{r}
library(tidyEmoji)
library(dplyr)

reviews <- data.frame(text = c("Best purchase ever \U0001f600\U0001f60d",
                               "It broke after a day \U0001f621",
                               "Does the job.",
                               "Wearing my mask \U0001f637\U0001f637",
                               "Shipped fast \U0001f3c1\U0001f600"))
```

### How much emoji is in the data?

```{r}
reviews %>% emoji_summary(text)        # entries with emoji vs. total
reviews %>% emoji_filter(text)         # keep only the rows that have emoji
```

### Which emoji are most common?

```{r}
reviews %>% emoji_frequency(text)      # every emoji, with name + category
reviews %>% top_n_emojis(text, n = 3)  # just the most frequent
```

### Pull the emoji out

`emoji_tokens()` gives one tidy row per emoji occurrence, with its name,
category and sentiment — ready to count, join or plot.

```{r}
reviews %>% emoji_tokens(text)
```

### Categorise and score sentiment

```{r}
reviews %>% emoji_categorize(text)     # which Unicode categories each row spans
reviews %>% emoji_sentiment(text)      # mean emoji sentiment per row (-1 to +1)
```

`emoji_sentiment()` uses the bundled **Emoji Sentiment Ranking** lexicon
(Kralj Novak et al., 2015). See the package vignette for a fuller tour.
