---
title: "Reversible emoji preprocessing for language models"
author: 'Youzhi Yu<br><span style="font-weight: 400; font-size: 0.85em;">University of Chicago</span>'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Reversible emoji preprocessing for language models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

## The decision nobody writes down

Every pipeline that sends user text to a language model makes a choice about
emoji, and most make it by accident: a regex in a cleaning script drops
anything non-ASCII, and the decision never appears in the methods section.

It is a consequential choice. An emoji is two or more tokens rather than one
character (measured below: 2 for a plain smiley, 13 for a family), models
disambiguate them poorly, and the glyphs readers disagree about are not the
obvious ones. In the bundled Emoji Sentiment Ranking, annotator disagreement
runs *against* the strength of the sentiment: `emoji_ambiguity()`'s entropy
correlates -0.31 with the absolute sentiment score, and averages 0.50 for the
strongly valenced glyphs against 0.76 for the middling ones. So a sentiment
score alone does not tell you which emoji are risky to strip.

Worse, most ways of removing emoji are one-way. If the pipeline has to hand
text back to a human, show a highlighted excerpt, or reconstruct what was
sent, a stripped emoji is gone.

`emoji_sanitize()` turns that accident into one named argument, and this
article is about the property the argument's values do *not* share:
**reversibility**. Only one policy both removes the emoji and survives a round
trip.

```{r setup}
library(tidyEmoji)
library(dplyr)
```

Everything here runs on the package's declared dependencies. There are no
optional packages and no network calls.

## A corpus built to break things

A round trip that works on a smiley proves very little. The cases that matter
are the multi-code-point ones: an emoji with a skin-tone modifier, a flag
built from two regional indicators, a family built from four people and three
joiners, a keycap built from a digit plus a presentation selector plus a
combining enclosing mark, and a glyph whose canonical spelling carries an
invisible presentation selector.

```{r}
awkward <- tibble::tibble(
  case = c("plain", "skin tone", "flag", "ZWJ family", "keycap",
           "bare heart", "no emoji"),
  text = c(
    "ship it \U0001f600 today",
    "nice work \U0001f44d\U0001f3fd",
    "landed in \U0001f1fa\U0001f1f8 already",
    paste("the whole",
          intToUtf8(c(0x1F468, 0x200D, 0x1F469, 0x200D,
                      0x1F467, 0x200D, 0x1F466)),
          "came"),
    paste("step", intToUtf8(c(0x31, 0xFE0F, 0x20E3)), "first"),
    paste("love it", intToUtf8(0x2764)),
    "nothing to see here"
  )
)

awkward %>%
  emoji_token_cost(text) %>%
  select(case, .emoji_n, .emoji_codepoints, .emoji_bytes,
         .emoji_token_estimate)
```

Note the third column. One visible emoji is not one code point, which is the
whole reason a hand-rolled substitution gets this wrong: the family is seven
code points and the keycap is three. The last column is why a prompt budget
notices: a plain smiley is estimated at two tokens and the family at thirteen.

## The ladder, run rather than described

The five policies form a ladder of information loss. `?emoji_sanitize`
tabulates it; here it is executed on one row, so you can see what a model
would actually receive.

```{r}
one <- awkward %>% filter(case == "ZWJ family")

for (p in c("keep", "shortcode", "name", "placeholder", "strip")) {
  cat(format(p, width = 12), "|",
      emoji_sanitize(one, text, policy = p)$text, "\n")
}
```

Now put them back. `text_to_emoji()` is the inverse of the shortcode form, and
the question is which policies it can recover from.

```{r}
restores <- function(policy, data = one) {
  sent <- emoji_sanitize(data, text, policy = policy)
  back <- text_to_emoji(sent, text)$text
  identical(back, data$text)
}

vapply(c("keep", "shortcode", "name", "placeholder", "strip"),
       restores, logical(1))
```

Two policies survive, and only one of them is a real preprocessing step:
`"keep"` changes nothing, so the interesting answer is `"shortcode"`. It is
the only policy that both removes the emoji from the model's input and lets
you reconstruct the user's text afterwards.

### It holds on the awkward cases

```{r}
awkward %>%
  mutate(
    sanitized = emoji_sanitize(., text, policy = "shortcode")$text,
    restored  = text_to_emoji(tibble::tibble(text = sanitized), text)$text,
    exact     = restored == text
  ) %>%
  select(case, sanitized, exact)
```

All seven come back byte for byte, but the last two rows are exact for a
reason worth separating out: nothing was substituted in them at all. The
`"no emoji"` row is obvious. The `"bare heart"` row is the interesting one,
because it looks like an emoji and is not detected as one.

```{r}
hearts <- tibble::tibble(
  spelling = c("bare U+2764", "qualified U+2764 U+FE0F"),
  text = c(paste("love it", intToUtf8(0x2764)),
           paste("love it", intToUtf8(c(0x2764, 0xFE0F))))
)

hearts %>% emoji_summary(text)
```

One of the two is counted. Detection asks for emoji presentation, and a bare
`U+2764` carries text presentation, so it passes through `emoji_sanitize()`
untouched and survives the round trip by never being touched. The fully
qualified spelling, which is what a keyboard emits, is detected and restores
byte for byte:

```{r}
sanitized_hearts <- emoji_sanitize(hearts, text, policy = "shortcode")
sanitized_hearts$text
text_to_emoji(sanitized_hearts, text)$text == hearts$text
```

This is a detection limitation rather than a reversibility one, and it is
described under *Detection* in `?tidyEmoji`. It matters here because text that
had its presentation selectors stripped upstream will quietly keep its emoji
through a policy that was supposed to remove them.

## `wrap` is part of the contract, not a style choice

`emoji_sanitize()` writes the shortcode as `:name:`, and `text_to_emoji()`
looks for exactly that. Decorate the token and the emoji still comes back, but
the decoration stays behind. Remove the colons and nothing comes back at all,
silently, because the shortcode is now an ordinary word.

```{r}
show_wrap <- function(w) {
  sent <- emoji_sanitize(one, text, policy = "shortcode", wrap = w)
  back <- text_to_emoji(sent, text)$text
  cat(format(w, width = 9), "|",
      if (identical(back, one$text)) "restored    " else "NOT restored",
      "|", back, "\n")
}

invisible(lapply(c(":{x}:", "[:{x}:]", "<{x}>", "::{x}::"), show_wrap))
```

The restored text in the last column separates the two failures. With a
decorating wrap the emoji does come back and the decoration is left stranded
around it. With `<{x}>` nothing comes back at all: the shortcode stays in the
text as an ordinary word. That third row is the one to watch for in a code
review, because nothing errors, the text looks cleaner, and the emoji are
unrecoverable.

## Do not throw away the signal you just removed

Sanitising for the model does not mean losing the emoji information. The point
of doing this in tidyEmoji rather than with a regex is that the same pipeline
can carry the emoji-derived features forward as ordinary columns, while the
text column that goes to the model holds no emoji at all.

```{r}
corpus <- tibble::as_tibble(
  utils::read.csv(
    system.file("extdata", "ata_tweets.csv", package = "tidyEmoji"),
    encoding = "UTF-8", stringsAsFactors = FALSE
  )
)

prepared <- corpus %>%
  emoji_sentiment(full_text) %>%          # how positive were the emoji
  emoji_risk(full_text) %>%               # how much annotators disagreed
  emoji_token_cost(full_text) %>%         # what they cost in the prompt
  mutate(prompt_text = emoji_sanitize(., full_text,
                                      policy = "shortcode")$full_text)

prepared %>%
  filter(.emoji_n > 0) %>%
  select(.emoji_n, .emoji_sentiment, .emoji_ambiguity_mean,
         .emoji_token_estimate, prompt_text) %>%
  head(3)
```

`prompt_text` is what you send. The four columns beside it are what you would
have destroyed, and they are the ones a model reads poorly anyway: a
disagreement score computed from how the lexicon's human annotators actually
split over each glyph is not something a prompt recovers.

## What it saves, and how to measure it honestly

```{r}
before <- corpus %>% emoji_token_cost(full_text)
after  <- prepared %>%
  select(full_text = prompt_text) %>%
  emoji_token_cost(full_text)

tibble::tibble(
  stage = c("original", "shortcode"),
  emoji = c(sum(before$.emoji_n), sum(after$.emoji_n)),
  emoji_bytes = c(sum(before$.emoji_bytes), sum(after$.emoji_bytes)),
  est_tokens = c(sum(before$.emoji_token_estimate),
                 sum(after$.emoji_token_estimate))
)
```

The sanitised column contains no emoji, so its emoji cost is zero by
construction. That is the saving, but treat the token figures as what they are
called: `.emoji_token_estimate` is an estimate. Pass your real tokeniser
through `emoji_token_cost(tokenizer = )` when the number goes in a budget,
and read `.emoji_bytes` and `.emoji_codepoints` when you want a fact rather
than a model of one.

## Which policy, by intent

| If the pipeline needs to | Use | Why |
|---|---|---|
| reconstruct the user's text afterwards | `"shortcode"` | the only reversible removal |
| feed a screen reader or a human reading a log | `"name"` | reads as words, not as a token |
| record that an emoji was present, not which | `"placeholder"` | keeps position, drops identity |
| discard emoji entirely and never look back | `"strip"` | leaves no trace, and no way back |
| leave the text alone | `"keep"` | the explicit no-op, visible in a diff |

The middle three are not lesser versions of `"shortcode"`. They answer
different questions, and each is the right answer to one of them. What matters
is that the choice is now written down in the script rather than implied by a
regex.

## Put the versions in the methods section

Which glyphs exist, what they are called, and which shortcode maps to which
emoji all depend on versions. A round trip is reproducible only against the
catalogue that performed it.

```{r}
emoji_provenance() %>% glimpse()
```

## See also

* `vignette("introduction", package = "tidyEmoji")` for the rest of the
  package: counting, categorising, sentiment and emotion scoring,
  co-occurrence, time series and model features.
* `?emoji_sanitize` for the reversibility table and the full `wrap` contract.
* `?emoji_token_cost` for what the estimate is and is not.
